Probabilistic Forecasts in Hierarchical~Time~Series

### 2018-07-06 ProbabilisticCoherence.tex
Cleanup of evaluation section

In the context of probabilistic forecast reconciliation there could be two motivations for using scoring rules.  The first is to compare unreconciled densities to reconciled densities.  Although reconciliation is a valuable goal in and of itself since it can be important in aligning decision making across, for example, different units of an enterprise, in the point forecasting literature, forecast reconciliation has also been shown to improve forecast performance \todo{Some References}.  It will be worthwhile to see whether the same holds in the probabilistic forecasting case.  The second motivation for using scoring rules is to compare two or more sets of reconciled probabilistic forecasts to one another.  The objective here is to evaluate which reconciliation mapping $g(.)$ works best in practice.

@@

One way to evaluate probabilistic forecasts is via the application of univariate scoring rules to each variable in a hierarchy.  A summary can be taken of the expected scores across each margin for example a mean or median.  In the simulations of section~\ref{sec:gaussian} we consider two scoring rules.  The log score is given by the log of the marginal density of each variable.  The cumulative rank probability score generalises mean square error and is given by
\begin{align} \label{eq:CRPS}
\text{CRPS}(\breve{F}_i,y_{T+h,i}) &=\int \left(\breve{F}_i(\breve{y}_i)-\mathbb{1}(y_i<y_{T+h,i})\right)dy_i\\ &=\E_{\breve{Y}_i}|\breve{Y}_{T+h,i}-y_{T+h,i}| - \frac{1}{2}\E_{\breve{Y}_i}|\breve{Y}_{T+h,i}-\breve{Y}^*_{T+h,i}|\,,
\end{align}
where $\breve{Y}_i$ and $\breve{Y}^*_{T+h,i}$ are independent copies of a random variable with distribution $\breve{F}_i$ and the latter is the predictive distribution for the $i^{th}$ margin of a forecast for time $t+h$ made at time $t$. The expectations in the second line can be approximated by Monte Carlo when a sample from the predictive distribution is available.

An advantage to this approach is that it allows the forecaster to evaluate the levels and individual series of the hierarchy where the gains to reconciliation are greatest.  For this reason this approach has been used in the limited literature on probabilistic forecasting for hierarchies \cite{BenTaieb2017} and Jeon et al \todo{Jeon Panagiotelis and Petropoulos reference} to date.  A major shortcomings to this approach however is that, evaluating univariate scores on the margins do not account for the dependence in the hierarchy.

@@

While the a number of alternative proper scoring rules are available for univariate forecasts, the multivariate case is somewhat more limited.  Here we focus on three scoring rules: the log score, the energy score and the variogram score.  These are summarized in table~\ref{table:scoringrules}.

\begin{table}[!bh]
  \caption{Scoring rules to evaluate multivariate forecast densities. Here, $\breve{\bm{y}}_{T+h}$ and $\breve{\bm{y}}^*_{T+h}$ are two independent random vectors from the coherent forecast distribution $\breve{\bm{F}}$ with density function $\breve{\bm{f}}(\cdot)$ at time $T+h$, and $\bm{y}_{T+h}$ is the vector of realizations. Further, $\breve{Y}_{T+h,i}$ and $\breve{Y}_{T+h,j}$ are the $i$th and $j$th components of the vector $\breve{\bm{Y}}_{T+h}$. The variogram score is given for order $p$, where $w_{ij}$ denote non-negative weights.}\label{table:scoringrules}
  \centering\small\setstretch{1.3}
  \begin{tabular}{@{}lp{8.1cm}l@{}}
    \toprule
    \textbf{Scoring rule}  & \textbf{Expression} & \textbf{Reference}           \\
    \midrule
    \text{Log score}       &
    $\text{LS}(\breve{\bm{F}},\bm{y}_{T+h}) = -\log {\breve{\bm{f}}(\bm{y}_{T+h})}$ &
    \citet{Gneiting2007}  \\\\[-0.2cm]
    \text{Energy score}    &
    $\text{ES}(\breve{\bm{Y}}_{T+h},\bm{y}_{T+h}) =
    \E_{\breve{\bm{F}}}
    \|\breve{\bm{Y}}_{T+h}-\bm{y}_{T+h}\|^\alpha -$ \par\hfill
    $\frac{1}{2}\E_{\breve{\bm{F}}}\|\breve{\bm{Y}}_{T+h}-\breve{\bm{Y}}^*_{T+h}\|^\alpha$, \,\, $\alpha \in (0,2]$ &
    \citet{Gneiting2008}  \\\\[-0.2cm]
    \text{Variogram score} &
    $\text{VS}(\breve{\bm{F}}, \bm{y}_{T+h}) =
    \sum\limits_{i=1}^{n}
    \sum\limits_{j=1}^{n}
    w_{ij}\Big(|y_{T+h,i} - y_{T+h,j}|^p -$ \par\hfill
    $\E_{\breve{\bm{F}}}|\breve{Y}_{T+h,i}-\breve{Y}_{T+h,j}|^p\Big)^2$     &
    \citet{SCHEUERER2015} \\
    \bottomrule
  \end{tabular}
\end{table}

The log score can be approximated using a sample of values from the probabilistic forecast density \citep{Jordan2017} however it is more commonly used when a parametric form for the density is available for the probabilistic forecast.  So far, we have mainly defined probabilistic forecasts in terms of probability measures.  Although densities can be obtained for both reconciled and unreconciled forecasts, the degeneracy of reconciled forecasts is problematic when using log scores. We will discuss this further in the next subsection.

The energy score on the other hand can be defined in terms of the characteristic function of the probabilistic forecast, but the representation in Table~\ref{table:scoringrules} in terms of expectations leads itself to easy computation when samples from the probabilistic forecast are available. 
An interesting case is where $\alpha=2$, where it can be easily shown that
\begin{equation} \label{eq:(5.1)}
\text{ES}(\breve{\bm{Y}_{T+h},y_{T+h}}) = \|\bm{y}_{T+h}-\breve{\bm{\mu}}_{T+h}\|^2,
\end{equation}
where $\breve{\bm{\mu}}_{T+h} =\E_{\bm{F}}(\breve{\bm{Y}}_{T+h}) $. In this limiting case, the energy score only measures the accuracy of the forecast mean, and not the entire distribution and the energy score is proper, but not strictly proper. \citet{Pinson2013a} also argues that the energy score has very low discriminative ability for incorrectly specified covariances, even though it discriminates the misspecified means well.

In contrast, \citet{SCHEUERER2015} have shown that the variogram score has a higher discrimination ability of misspecified means, variances and correlation structures than the energy score. For a finite sample of size $B$ from the multivariate forecast density $\breve{\bm{F}}$, the empirical variogram score is defined as
\begin{equation}
\text{VS}(\breve{\bm{F}}, \bm{y}_{T+h}) = \displaystyle\sum_{i=1}^{n}\displaystyle\sum_{j=1}^{n}w_{ij}\left(|y_{T+h,i} - y_{T+h,j}|^p - \frac{1}{B} \displaystyle\sum_{k=1}^{B} |\breve{Y}^k_{T+h,i}-\breve{Y}^k_{T+h,j}|^p\right)^2.
\end{equation}
\citet{SCHEUERER2015} recommend using $p=0.5$.

@@

\begin{theo}[Impropriety of log score]
	When the true data generating process is a coherent measure, then the log score is improper with respect to the class of incoherent measures.
\end{theo}
\begin{proof}
Consider a rotated version of hierarchical time series $\bm{z}_t={\bm U}{\bm y}_t$ so that the first $m$ elements of $\bm{z}_t$ denoted $\bm{z}^{(1)}_t$ are unconstrained, while the remaining $n-m$ elements denoted $\bm{z}^{(2)}_t$ equal $0$ when the aggregation constraints hold. An example of the $n\times n$ ${\bm U}$ could be the matrix of left singular vectors of ${\bm S}$.  For a non-degenerate probability measure on $\mathbb{R}^n$, the density is the Radon-Nikodym derivative with respect to the usual Lesbegue measure on $\mathbb{R}^n$.  %For a coherent probability measure the density after rotation is the Radon-Nikodym derivative with respect to the usual Lesbegue measure on $\mathbb{R}^m$.

Consider the case where the true density is $f_1(\bm{z}^{(1)}_t)\mathbb{1}\left({\bm{z}^{(2)}_t}\right)$, and is compared to an incoherent density is given by $f_1(\bm{z}^{(1)}_t)f_2(\bm{z}^{(2)}_t)$, where $f_2$ is highly concentrated around $0$ but still a proper density. For example $f_2$ may be Gaussian with variance $\sigma^2{\bm I}$ with $\sigma^2 < (2\pi)^{-1}$. The log score under the true DGP is
$$
S\left(\tilde{f},\bm{z}^{(1)}_t\right) = -\log f_1\left(\bm{z}^{(1)}_t\right),
$$
while that of the unreconciled density is
\begin{align}
S\left(\hat{f},\bm{z}^{(1)}_t\right) &= -\log f_1(\bm{z}^{(1)}_t)-f_2(\bm{z}^{(1)}_t)\\
&= -\log f_1(\bm{z}^{(1)}_t)+\frac{n-m}{2}\log(2\pi\sigma^2)\\
&<-\log f_1(\bm{z}^{(1)}_t).
\end{align}

After taking expectations $ES(f,f) > ES(\hat{f},f)$, violating the condition \eqref{eq:(3.1.)} for a proper scoring rule.  
\end{proof}

A similar issue also arises when discrete random variables are modelled as if they were continuous, an issue discussed in Section 4.1, page 366 of \cite{Gneiting2007}. This implies that the log score should not be used to evaluate multivariate densities with degeneracies and should be avoided when comparing reconciled and unreconciled probabilistic forecasts.

@@

Coherent probabilistic forecasts can be completely characterised in terms of basis series; if a probabilistic forecast is  available for the basis series then a probabilistic forecast can be recovered for  the  entire hierarchy via Definition~\ref{def:cohprob}. This may suggest that it is adequate to merely compare two coherent forecasts to  one  another using  the basis series only.  We now show how this is  dependent on the specific scoring rule used.

For the log score, suppose the coherent probabilistic forecast has density $f(\bm{b})$.  The density for the full hierarchy is given by $f(\bm{y})=f(\bm{Sb})=f(\bm{b})J^{-1}$ where $J=\prod_{j=1}^{m}\lambda_j$ is a pseudo-determinant of the non-square  matrix ${\bm S}$ and $\lambda_j$ are the non-zero singular values of ${\bm S}$.  Therefore for any coherent density the log score of the full hierarchy differs from the log score for the bottom level series by the term $log(J)$.  This term depends  only on the structure of the  hierarchy and is fixed across different reconciliation methods.  Therefore if one method achieves a lower expected log score compared to an alternative method  when assessed using the bottom level series, the same ordering is preserved when an assessment is made on the basis of the full hierarchy.

@@

The properties of multivariate scoring rules in the context of evaluating reconciled probabilistic forecasts in Table~\ref{tab:prop}.

>>>>

In the context of probabilistic forecast reconciliation there could be two motivations for using scoring rules.  The first is to compare unreconciled densities to reconciled densities.  Although reconciliation is a valuable goal in and of itself since it can be important in aligning decision making across, for example, different units of an enterprise, in the point forecasting literature, forecast reconciliation has also been shown to improve forecast performance \citep{AthEtAl2017, Wickramasuriya2017}.  It will be worthwhile to see whether the same holds in the probabilistic forecasting case.  The second motivation for using scoring rules is to compare two or more sets of reconciled probabilistic forecasts to one another.  The objective here is to evaluate which reconciliation mapping $g(.)$ works best in practice.

@@

One way to evaluate probabilistic forecasts is via the application of univariate scoring rules to each variable in a hierarchy.  A summary can be taken of the expected scores across each margin for example a mean or median.  In the simulations of section~\ref{sec:gaussian} we consider two scoring rules.  The log score is given by the log density, in this case for each margin of the probabilistic forecast.  The cumulative rank probability score generalises mean square error and is given by
\begin{align} \label{eq:CRPS}
\text{CRPS}(\breve{F}_i,y_{i}) &=\int \left(\breve{F}_i(\breve{Y}_i)-\mathbb{1}(y_i<y_{i})\right)d\breve{Y}_i\\ &=\E_{\breve{Y}_i}|\breve{Y}_{i}-y_{i}| - \frac{1}{2}\E_{\breve{Y}_i}|\breve{Y}_{i}-\breve{Y}^*_{i}|\,,
\end{align}
where $\breve{F}_i$ is the cumulative distribution function of the $i^{th}$ margin of the probabilistic forecast, $\breve{Y}_i$ and $\breve{Y}^*_{T+h,i}$ are independent copies of a random variable with distribution $\breve{F}_i$ and $y_i$ is the outcome of the $i^{th}$ margin. The expectations in the second line can be approximated by Monte Carlo when a sample from the predictive distribution is available.

An advantage to this approach is that it allows the forecaster to evaluate the levels and individual series of the hierarchy where the gains from reconciliation are greatest.  For this reason this approach has been used in the limited literature on probabilistic forecasting for hierarchies \citep{BenTaieb2017, JeoEtAl2018} to date.  A major shortcomings of this approach however, is that evaluating univariate scores on the margins do not account for the dependence in the hierarchy.

@@

While a number of alternative proper scoring rules are available for univariate forecasts, the multivariate case is somewhat more limited.  Here we focus on three scoring rules: the log score, the energy score and the variogram score.  %These are summarized in table~\ref{table:scoringrules}.
%
%\begin{table}[!bh]
%  \caption{Scoring rules to evaluate multivariate forecast densities. Here, $\breve{\bm{y}}_{T+h}$ and $\breve{\bm{y}}^*_{T+h}$ are two independent random vectors from the coherent forecast distribution $\breve{\bm{F}}$ with density function $\breve{\bm{f}}(\cdot)$ at time $T+h$, and $\bm{y}_{T+h}$ is the vector of realizations. Further, $\breve{Y}_{T+h,i}$ and $\breve{Y}_{T+h,j}$ are the $i$th and $j$th components of the vector $\breve{\bm{Y}}_{T+h}$. The variogram score is given for order $p$, where $w_{ij}$ denote non-negative weights.}\label{table:scoringrules}
%  \centering\small\setstretch{1.3}
%  \begin{tabular}{@{}lp{8.1cm}l@{}}
%    \toprule
%    \textbf{Scoring rule}  & \textbf{Expression} & \textbf{Reference}           \\
%    \midrule
%    \text{Log score}       &
%    $\text{LS}(\breve{\bm{F}},\bm{y}_{T+h}) = -\log {\breve{\bm{f}}(\bm{y}_{T+h})}$ &
%    \citet{Gneiting2007}  \\\\[-0.2cm]
%    \text{Energy score}    &
%    $\text{ES}(\breve{\bm{Y}}_{T+h},\bm{y}_{T+h}) =
%    \E_{\breve{\bm{F}}}
%    \|\breve{\bm{Y}}_{T+h}-\bm{y}_{T+h}\|^\alpha -$ \par\hfill
%    $\frac{1}{2}\E_{\breve{\bm{F}}}\|\breve{\bm{Y}}_{T+h}-\breve{\bm{Y}}^*_{T+h}\|^\alpha$, \,\, $\alpha \in (0,2]$ &
%    \citet{Gneiting2008}  \\\\[-0.2cm]
%    \text{Variogram score} &
%    $\text{VS}(\breve{\bm{F}}, \bm{y}_{T+h}) =
%    \sum\limits_{i=1}^{n}
%    \sum\limits_{j=1}^{n}
%    w_{ij}\Big(|y_{T+h,i} - y_{T+h,j}|^p -$ \par\hfill
%    $\E_{\breve{\bm{F}}}|\breve{Y}_{T+h,i}-\breve{Y}_{T+h,j}|^p\Big)^2$     &
%    \citet{SCHEUERER2015} \\
%    \bottomrule
%  \end{tabular}
%\end{table}

The log score can be approximated using a sample of values from the probabilistic forecast density \citep{Jordan2017} however it is more commonly used when a parametric form for the density is available for the probabilistic forecast.  The energy score on the other hand can be defined in terms of the characteristic function of the probabilistic forecast, but the representation in terms of expectations leads itself to easy computation when samples from the probabilistic forecast are available. 
An interesting limiting case is where $\alpha=2$, where it can be easily shown that
\begin{equation} \label{eq:(5.1)}
\text{ES}(\breve{\bm{Y}_{T+h},y_{T+h}}) = \|\bm{y}_{T+h}-\breve{\bm{\mu}}_{T+h}\|^2,
\end{equation}
where $\breve{\bm{\mu}}_{T+h} =\E_{\bm{F}}(\breve{\bm{Y}}_{T+h}) $. In this limiting case, the energy score only measures the accuracy of the forecast mean, and not the entire distribution and the energy score is proper, but not strictly proper. \citet{Pinson2013a} also argues that the energy score has very low discriminative ability for incorrectly specified covariances, even though it discriminates the misspecified means well.

In contrast, \citet{SCHEUERER2015} have shown that the variogram score has a higher discrimination ability of misspecified means, variances and correlation structures than the energy score. When $\breve{\bm Y}$ is a random variable from probabilistic forecast $\breve{F}$, the empirical variogram score is defined as
\begin{equation}
\text{VS}(\breve{F}, \bm{y}) = \displaystyle\sum_{i=1}^{n}\displaystyle\sum_{j=1}^{n}w_{ij}\left(|y_{i} - y_{j}|^p - E_{\breve{Y}_i,\breve{Y}_j} |\breve{Y}^k_{i}-\breve{Y}^k_{j}|^p\right)^2.
\end{equation}
\citet{SCHEUERER2015} recommend using $p=0.5$.

@@

\begin{theo}[Impropriety of log score]
	When the true data generating process is coherent, then the log score is improper with respect to the class of incoherent measures.
\end{theo}
\begin{proof}
Consider a rotated version of hierarchical time series $\bm{z}_t={\bm U}{\bm y}_t$ so that the first $m$ elements of $\bm{z}_t$ denoted $\bm{z}^{(1)}_t$ are unconstrained, while the remaining $n-m$ elements denoted $\bm{z}^{(2)}_t$ equal $0$ when the aggregation constraints hold. An example of the $n\times n$ ${\bm U}$ could be the matrix of left singular vectors of ${\bm S}$.  %For a non-degenerate probability measure on $\mathbb{R}^n$, the density is the Radon-Nikodym derivative with respect to the usual Lesbegue measure on $\mathbb{R}^n$.  %For a coherent probability measure the density after rotation is the Radon-Nikodym derivative with respect to the usual Lesbegue measure on $\mathbb{R}^m$.

Consider the case where the true predictive density is $f_1({\bm z}^{(1)}_t)\mathbb{1}\left({\bm z}^{(2)}_t={\bm 0}\right)$, and we evaluate an incoherent density is given by $f_1({\bm z}^{(1)}_t)f_2({\bm z}^{(2)}_t)$, where $f_2$ is highly concentrated around $0$ but still a non-degenerate. For example $f_2$ may be Gaussian with variance $\sigma^2{\bm I}$ with $\sigma^2 < (2\pi)^{-1}$. The log score under the true data generating process is
$$
S\left(\tilde{f},{\bm z}^{(1)}_t\right) = -\log f_1\left({\bm z}^{(1)}_t\right),
$$
while that of the unreconciled density is
\begin{align}
S\left(\hat{f},\bm{z}^{(1)}_t\right) &= -\log f_1(\bm{z}^{(1)}_t)-f_2(\bm{z}^{(1)}_t)\\
&= -\log f_1(\bm{z}^{(1)}_t)+\frac{n-m}{2}\log(2\pi\sigma^2)\\
&<-\log f_1(\bm{z}^{(1)}_t).
\end{align}

After taking expectations $ES(f,f) > ES(\hat{f},f)$, violating the condition in Equation~\eqref{eq:(3.1.)} for a proper scoring rule.  
\end{proof}

A similar issue also arises when discrete random variables are modelled as if they were continuous, an issue discussed in Section 4.1 of \cite{Gneiting2007}. This implies that the log score should be avoided when comparing reconciled and unreconciled probabilistic forecasts.

@@

Coherent probabilistic forecasts can be completely characterised in terms of basis series; if a probabilistic forecast is  available for the basis series then a probabilistic forecast can be recovered for  the  entire hierarchy via Definition~\ref{def:cohprob}. This may suggest that it is adequate to merely compare two coherent forecasts to  one  another using  the basis series only.  We now show how this  depends on the specific scoring rule used.

For the log score, suppose the coherent probabilistic forecast has density $f(\bm{b})$.  The density for the full hierarchy is given by $f(\bm{y})=f(\bm{Sb})=f(\bm{b})J^{-1}$ where $J=\prod_{j=1}^{m}\lambda_j$ is a pseudo-determinant of the non-square  matrix ${\bm S}$ and $\lambda_j$ are the non-zero singular values of ${\bm S}$.  Therefore for any coherent density the log score of the full hierarchy differs from the log score for the bottom level series by the term $log(J)$.  This term depends  only on the structure of the  hierarchy and is fixed across different reconciliation methods.  Therefore if one method achieves a lower expected log score compared to an alternative method  using the bottom level series only, the same ordering is preserved when an assessment is made on the basis of the full hierarchy.

@@

The properties of multivariate scoring rules in the context of evaluating reconciled probabilistic forecasts are summarised in Table~\ref{tab:prop}.
